Towards a classification framework for social machines

### 2013-02-20 latex/paper_new.tex
New structure of the paper
Following discussions with Wendy

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{Social Machines Workshop}{2013, Paris, France}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{A Framework for Classification of Social Machines}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\section{Introduction (0.5 page)}

What are social machines and why are they important

An emerging, interdisciplinary field of research and development with a growing number of social machine instances and studies about them

We propose a classification framework for social machines. Targeted audiences

@@

\section{Methodology (1 page)}

%How was the classification framework defined
%
%* interviews with XXX researchers on key descriptive features of systems they are familiar with
%
%* consolidation of features listed for a resulting set of 50 (?) social machines
%
%* grid analysis to identify groups of similar systems
%
%* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems
%

In order to identify and classify social machines, we first embarked on a principled
knowledge elicitation process~\cite{knowledgeelicitation}, to determine the {\it elements}
(i.e., the social machines) and the {\it constructs} (i.e., the classifying factors) that
we can use to classify social machines. We used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of elements and constructs. We
asked ten Computer Science researchers familiar with social machines to create their own
repertory grids, and populate the elements from their own knowledge, and to create the
constructs using the standard repertory grid elicitation technique. This exercise led
to 10 grids, the union of which comprised a total of 56 unique elements (social machines)
and 117 different constructs.

@@

\section{Classification framework (2 pages)}

%First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc

To create a classification framework, it is necessary to first determine the aspects of
classification. We will first discuss the relationships between social
machines, enabling technologies and platforms; we move on to discuss the multi-faceted
nature of motivation within social machines; next we evaluate the artefacts of social
machines: contributions, actions, activities and purpose; finally we discuss the metrics of
participation, and the difference between designed and emergent properties of social
machines.

\subsection{The Hierarchical/Polyarchical relationship of social machines}

While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
social machines to be created, and instantiations of those frameworks into social machines that have been created for a specific purpose, such as Wikipedia
and Ushahidi Haiti. However, through further investigation, we discovered that the distinction between framework and instantiations was not the only
differentiator. Specifically, that some social machines could also have sub-machines or communities within them, that are working to solve different problems,
within the context of the larger machine. For example, Wikipedia is an instantiation of the MediaWiki framework, but there are a large number of
communities within Wikipedia that work on domain-specific areas, such as those contributing in areas of science and technology\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}. However, we must also note that wikipedia is primarily split into communities of language, with the English, German and French wikipedias having the largest number of articles. It is therefore possible to analyse wikipedia into a hierarchy, from MediaWiki, to the languages of Wikipedia, down to the domain-specific communities. We explored this idea further, to include other social machines, and links where social machines utilise other social machines in order to function. For example, the social machine used by the Obama '12 US Presidential campaign~\cite{obamakieron}, which relied on social machines such as Twitter and Facebook.

@@

\subsection{Multi-faceted motivation, and who-gets-what benefit}

@@

\subsection{Artefacts of Social Machines}

@@

\subsection{Metrics of participants and usage}

@@

\section{Usage examples (0.5 page)}

Wikipedia

GalaxyZoo (?)

Ushahidi

Facebook

MechanicalTurk

GWAPs

\subsection{Causes as Social Machines}

[more about this?]

\section{Related work (0.5 pages)}

existing classification systems in CSCW, collective intelligence, social computing systems

\section{Conclusions and future work (0.5 pages)}

Detail evaluation plans

Feedback from different disciplines, and from the developers community

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

>>>>

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{Workshop on the theory and practice of social machines @ WWW2013}{2013, Rio de Janeiro, Brazil}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Towards a classification framework for social machines}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\section{Social machines: an introduction}
Once upon a time 'machines' were programmed by programmers and used by users. The Web, and the
massive participation in it, has changed this relationship: we now see configurations of people interacting with content exposed via the Web and with each other, typified by social Web sites. Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'  ñ  a  machine  comprising  digital  and  human
components and the interactions between them. This crucial transition in thinking acknowledges the reality of today's computing systems and is essential to underpin any understanding of the science and engineering of our future socio-technical systems: we view an ecosystem of co-evolving social machines.

The  typical  social Web  site  is  a  social machine, with  its user-generated  content  and  social network evolving dynamically  and  hand-in-hand. More  than  just  'data' or  'content',  these sites have  some notion of  'social artefacts' around which we forge the links in the social network, like photos on flickr or movies on YouTube, and they  can also be  flourishing  spaces  for human  ingenuity.  
Today we  see  a  divide  between  conventional  computer  systems  dedicated  to  data-  and  computation-intensive tasks, and social computing like the Web 2.0 sites. But many of our challenges, from healthcare to transport, need  both  of  these  axes:  a  sophisticated  combination  of  data-intensive  automation  and  deep  social involvement. This suggests the need for new social machines to tackle new problems, and to achieve this we
need to understand the science and engineering of social machines.

A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for content, has some of the right ingredients but there is really no social component in this strict provider-consumer relationship. When we fill out a form (e.g. health information, birds spotted in my garden) it  is minimally  a  social machine  because humans  are  part  of  the  'social  computation'.  This  becomes  richer when  the database  is  curated by members  of  the  community  (e.g. Wikipedia) and when  the  social network adds value implicitly (amazon) or explicitly (Facebook). We anticipate many categories of social machine in this continuum, and they may manifest themselves as Web site or app, or indeed in the Internet of Things.

An emerging, interdisciplinary field of research and development with a growing number of social machine instances and studies about them.

We propose a classification framework for social machines.

Targeted audiences

@@

Dimensions of the classification

* tasks and types of contributions: what does the system do, how do participants contribute: from supplying data and creating content to solving problems or validating solutions.

* social network and interaction: roles, hierarchies of roles, profiling

* motivation and incentives: system features influencing user behavior and the social network dynamics

* human-machine interaction: how are human contributions used to compute things, including quality control, spam, workflow, aggregation of results

People  supply  or  refine  data.    Systems  in  this  category  use  people  as,  effectively,  sophisticated  sensors.
Examples of  this  sort of  system  are Galaxy  Zoo  [www.galaxyzoo.org] where people  classify  galaxies with
collective performance as good as professional astronomers and reCaptcha [www.google.com/recaptcha]
where people refine data describing digital manuscripts.
Åú People are elementary problem solvers. Systems in this category typically find innovative ways of getting
people to perform a task that are easy for them but for which no effective computer algorithm is available.
Examples  of  this  sort  of  system  are Mechanical  Turk  [www.mturk.com]  and  the  related  Turkit  system
[groups.csail.mit.edu/uid/turkit]  that  allows  these  actions  to  be  strung  together  iteratively,  thus  giving  a
basic form of composition.
Åú People generate/test partial  solutions. Systems  in  this  category provide  incentives  for people  to perform
complex  problem  solving  that  generates  a  design  to  be  tested  automatically  and/or  tests  automatically
generated  designs.    Examples  include  EteRNA  [eterna.cmu.edu]  in which  participants  design RNA  that  is
evaluated automatically using simulators and then refined via human effort; Matlab programming contests
[www.mathworks.co.uk/matlabcentral/contest] where people compete for the best solution as judged by automated  tests  that  compute  the  efficiency  of  the  code;  or  Scratch  [scratch.mit.edu]  in  which  simple programs are created by some participants and modified by others in a remixing process.

\section{Comparison with related fields}

Web 2.0

Collective Intelligence, Wisdom of the crowds

Open Innovation

In  social machines  the  social and  the  technical are equal  partners. While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the
social machine perspective is one of incentivised participation and not humans enslaved to computers.

Crowdsourcing, human computation

CSCW, social computing

\section{Classification of social machines}

\subsection{Methodology}
In order to identify and classify social machines, we first embarked on a principled
knowledge elicitation process~\cite{knowledgeelicitation}, to determine the {\it elements}
(i.e., the social machines) and the {\it constructs} (i.e., the classifying factors) that
we can use to classify social machines. We used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of elements and constructs. We
asked ten Computer Science researchers familiar with social machines to create their own
repertory grids, and populate the elements from their own knowledge, and to create the
constructs using the standard repertory grid elicitation technique. This exercise led
to 10 grids, the union of which comprised a total of 56 unique elements (social machines)
and 117 different constructs.

@@

\subsection{Classification dimensions}

\subsubsection{The polyarchical relationship of social machines}
While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
social machines to be created, and instantiations of those frameworks into social machines that have been created for a specific purpose, such as Wikipedia
and Ushahidi Haiti. However, through further investigation, we discovered that the distinction between framework and instantiations was not the only
differentiator. Specifically, that some social machines could also have sub-machines or communities within them, that are working to solve different problems,
within the context of the larger machine. For example, Wikipedia is an instantiation of the MediaWiki framework, but there are a large number of
communities within Wikipedia that work on domain-specific areas, such as those contributing in areas of science and technology\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}. However, we must also note that wikipedia is primarily split into communities of language, with the English, German and French wikipedias having the largest number of articles. It is therefore possible to analyse wikipedia into a hierarchy, from MediaWiki, to the languages of Wikipedia, down to the domain-specific communities. We explored this idea further, to include other social machines, and links where social machines utilise other social machines in order to function. For example, the social machine used by the Obama '12 US Presidential campaign~\cite{obamakieron}, which relied on social machines such as Twitter and Facebook.

@@

\subsubsection{Multi-faceted motivation, and who-gets-what benefit}

@@

\subsubsection{Artefacts of Social Machines}

@@

\subsubsection{Metrics of participants and usage}

@@

\subsection{Using the classification framework}

In order to demonstrate usage of our framework's constructs, we ranked our social machines
by their Alexa ranking, and performed a full repertory grid elicitation exercise. The
results are illustrated in a grid and accompanying dendrograms for element and construct
similarity in Figure~\ref{dendrogram}.

\begin{figure*}
\begin{center}
\includegraphics[width=18cm]{img/dendrogram.png}
\caption{Dendrogram from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
\end{center}
\end{figure*}

[needs some discussion on what this shows.]

\section{Future work: building a social machine around the classification framework}

\section{Acknowledgments}
